### 2023-04-11 reachable-resilience.tex
up

@@

\problemx{$t$-liveness}
{Vass $V=(Q,T)$, $t \in T$}
{Is $t$ live ?}

@@

The characterization is used to show that VASS are $\uparrow \post^*$-effective. Since the downward-reachability problem is decidable for VASS \alain{ref ?}, we \alain{we ou Ozkan ? qui a fait la preuve ?} deduce the following:

\begin{proposition}
Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.\alain{où sont définis les d-VASS, $V= (Q,T)$, $q(\textbf{v}) \in Q \times \N^d$ ?}
\end{proposition}
%\alain{mauvaises notations, la tradition est que P est l'ensemble des places...les états/markings en minuscules,....}

@@

\begin{proof}
Consider Reset $d$-VASS
in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} \in \mathbb{N}^d$ 
%			contains only the minimal element $\min(S) = 0$ %$0$ 
\alain{tu n'as pas le doit d'écrire $0$ sans explication}
and $\Bad = \mathbb{N}^d \setminus \Safe$, 
it follows the state-resilience problem \alain{écris toujours pareil les expressions choisies comme state-resilience}is equivalent 
to the problem of whether $0$ can be visited infinitely often. \alain{incompréhensible, explique}
% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
Hence the undecidability.
\end{proof}

@@

\begin{theorem}\label{downward srp}
{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,\alain{a-t-on aussi la upward compatibility ?}
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
\end{theorem}

\begin{proof}
In order to decide whether state-resilience hold, we execute two procedures in parallel,
one looking for a resilience certificate and one looking for a non-resilience certificate.
Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward-closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
Notice that, by standard \alain{upward ? downward ?}compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.

>>>>

\mathieu{En train de penser que le mieux serait de définir la $t$-liveness pas pour les WSTS mais pour les (Reset) VASS uniquement et ensuite dire SRP indécidable dans WSTS car indécidable pour les Reset VASS.}

@@

\problemx{(Reset) VASS $t$-liveness}
{(Reset) VASS $V=(Q,T)$, $t \in T$}
{Is $t$ live ?}

@@

The characterization is used to show that VASS are $\uparrow \post^*$-effective. Since the downward-reachability problem is decidable for VASS \alain{ref ?}, we \alain{we ou Ozkan ? qui a fait la preuve ?} 
\mathieu{le papier d'Ozkan: 
[Petri nets (equipped with any
finite set of initial states) are $\uparrow \post^*$-effective by above Prop.: Reachability for Petri nets is decidable and recursively equivalent to submarking reachability.
This corresponds to the anti-ideal reachability problem for Petri nets.]}
deduce the following:

\begin{proposition}
Let $V= (Q,T)$ be a $d$-VASS and $q(\textbf{v}) \in Q \times \N^d$ a configuration, then one can compute a basis of $\uparrow \post^*(q(\textbf{v}))$.\alain{où sont définis les d-VASS, $V= (Q,T)$, $q(\textbf{v}) \in Q \times \N^d$ ?}
\end{proposition}

@@

\begin{proof}
Consider Reset $d$-VASS
in the case $\Safe=\{ \textbf{0} \}$ where  $ \textbf{0} = 0^d \in \mathbb{N}^d$ 
%			contains only the minimal element $\min(S) = 0$ %$0$ 
\alain{tu n'as pas le doit d'écrire $0$ sans explication}
and $\Bad = \mathbb{N}^d \setminus \Safe$, 
it follows the 
state-resilience problem \alain{écris toujours pareil les expressions choisies comme state-resilience}
is equivalent 
to the problem of whether $0$ can be visited infinitely often. \alain{incompréhensible, explique}
% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
Hence the undecidability.
\end{proof}

@@

\begin{theorem}\label{downward srp}
{\sc state-resilience} is decidable for ideal effective WSTS with downward compatibility,\alain{a-t-on aussi la upward compatibility ?}
\mathieu{oui}
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
\end{theorem}

\begin{proof}
In order to decide whether state-resilience hold, we execute two procedures in parallel,
one looking for a resilience certificate and one looking for a non-resilience certificate.
Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward-closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
Notice that, by 
%standard \alain{upward ? downward ?}
upward compatibility, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.

### 2023-04-11 resilience-safe-down.tex
Up
up

\begin{proof}
By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
%  Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\pred^*(\Safe)$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
The resilience problem can be reformulated as 
% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$.
$\Bad \subseteq  \pred^*(\Safe)$.
Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of closed \alain{non définis} subsets.
Hence we can compute the intersection of
% $\uparrow \{b_1,b_2,...,b_m\}$
$\Bad$
and
$S \setminus \pred^*(\Safe)$,
which are both upward-closed.
Since
% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$
$\Bad \subseteq \pred^*(\Safe)$
can be reformulated as
$\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
the resilience problem is decidable.
\end{proof}

@@

Let $D$ be a downward-closed subset of $S$
and let $x \in \downarrow \pred^*(D)$.
By design \alain{?} there exists
$y \in \pred^*(D)$ 
such that $x \leq y$
and $d \in D$ such that
$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.

By downward compatibility $a_0 \to a_1$
implies there exists $a'_1 \leq a_1$ with
$x \to^* a'_1$.
More generally $a_i \to a_{i+1}$ and
$a'_i\leq a_i$ implies the existence of $a'_{i+1} \leq a_{i+1}$ such that
$a'_i \to^* a'_{i+1}$,
and, by induction,
 $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* d'$
where
$d' \leq d$ \alain{qui appartient à qui ?}belongs to $D$ by downward closure of $D$.
Hence $x \in \pred^*(D)$.
\end{proof}

@@

\alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}

and $\Bad = \uparrow U \setminus \Safe $

@@

\begin{proof}
Since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong %upward-
compatibility, then $\pred^n(\Safe)= \uparrow~\pred^n(\Safe)$ for all $n \in \N$,\alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
and there exists $n_0 \in \N$ such that 
$\pred^{n_0}(\Safe) = \uparrow \pred^{n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
Hence the equivalence.
\end{proof}

@@

\begin{proof}
Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WTSTS with strong (upward) compatibility,  $\pred^n(\Safe)=\uparrow \pred^n(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$. % Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm. 
Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$. % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
hence the resilience problem is decidable.
\end{proof}

>>>>

\begin{proof}
By hypothesis $\pred^*(\Safe)$ is downward-closed, since $\Safe$ is downward-closed.
%  Let $\{J_1, J_2,...,J_n\}$ be the ideal decomposition of $\pred^*(\Safe)$ and $\{b_1,b_2,...,b_m\}$ be the (unique) minimal basis of $\Bad$.
The resilience problem can be reformulated as 
% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$.
$\Bad \subseteq  \pred^*(\Safe)$.
Since $\mathscr{S}=(S,\rightarrow, \leq)$ is ideally effective, we can compute intersections of upward- or downward-closed 
%\alain{non définis}
 subsets.
Hence we can compute the intersection of
% $\uparrow \{b_1,b_2,...,b_m\}$
$\Bad$
and
$S \setminus \pred^*(\Safe)$,
which are both upward-closed.
Since
% $\uparrow \{b_1,b_2,...,b_m\} \subseteq J_1 \cup J_2 \cup \cdots \cup J_n$
$\Bad \subseteq \pred^*(\Safe)$
can be reformulated as
$\Bad \cap (S \setminus \pred^*(\Safe)) = \emptyset$
the resilience problem is decidable.
\end{proof}

@@

Let $D$ be a downward-closed subset of $S$
and let $x \in \downarrow \pred^*(D)$.
By downward closure, there exists
$y \in \pred^*(D)$ 
such that $x \leq y$.
By definition of $\pred^*(D)$ then there exists 
$d \in D$, $m\geq 0$ and $(a_i)_{0 \leq i \leq m+1} \in S^{m+2}$ such that
$y = a_0 \to a_1 \to a_2 \to \cdots \to a_m \to a_{m+1} = d$.

By downward compatibility $a_0 \to a_1$
implies that there exists $a'_1 \in S$ such that $a'_1 \leq a_1$ and
$x \to^* a'_1$.
More generally $a_i \to a_{i+1}$ and
$a'_i\leq a_i$ implies the existence of $a'_{i+1} \in S$ with $a'_{i+1} \leq a_{i+1}$ and
$a'_i \to^* a'_{i+1}$,
and, by induction,
 $x \to^* a'_1 \to^* \cdots \to^* a'_{m} \to^* a'_{m+1} = d'$
with $d' \leq d$.
Since
$d'$ 
% \alain{qui appartient à qui ?}
belongs to $D$ by downward closure of $D$, $x \in \pred^*(D)$.
\end{proof}

@@

\alain{à quoi ça sert ici ?: quelle signification ?Decidability of $U$ and $\Safe$ imply the decidability of $\Bad$.}
\mathieu{Pas convaincu que ça serve, je vais enlever}

@@

\begin{proof}
Since $\Safe=\uparrow \Safe$ and
$\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong %upward-
compatibility, then $\pred^{\leq n}(\Safe)= \uparrow~\pred^{\leq n}(\Safe)$ for all $n \in \N$,\alain{$\pred^n$ ou $\pred^{\leq n}$ ? même question pour post. et où est-ce défini ?}
and there exists $n_0 \in \N$ such that 
$\pred^{\leq n_0}(\Safe) = \uparrow \pred^{\leq n_0}(\Safe) = \uparrow \pred^*(\Safe) = \pred^*(\Safe)$.
Hence the equivalence.
\end{proof}

@@

\begin{proof}
Since $\Safe=\uparrow \Safe$ is upward-closed, there exists a finite basis $B_{ \Safe}$ such that $\uparrow \Safe = \uparrow B_{\Safe}$. 
Moreover since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS with strong (upward) compatibility,  $\pred^{\leq n}(\Safe)=\uparrow \pred^{\leq n}(\Safe)$ for all $n\in \N$ and $\pred^{\leq k}(\Safe)$ admits a finite basis $B_{\pred^{\leq k}(\Safe)}$. % Since $\mathscr{S}=(S,\rightarrow,\leq)$ is a WSTS  with effective pred-basis, we may compute a finite basis of $\pred^{\leq k}(\Safe)$ with the backward coverability algorithm. 
Since $U$  is upward-closed, there exists a finite basis $B_{U}$ such that $U = \uparrow B_{U}$. % Moreover $ \uparrow B_{\Bad} \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{\Bad}$, there is a $s \in B_{\pred^*(\Safe)}$ such that $s \leq b$.
Since $\Bad = U \setminus \Safe$, $\Bad \subseteq \uparrow B_{\pred^{\leq k}(\Safe)}$ iff for every $b \in B_{U}$, there is either $s \in B_{\pred^{\leq k}(\Safe)}$ such that $s \leq b$, either $t \in B_{\Safe}$ such that $t\leq b$,
% si on est dans up Bad, il faut que ou bien on soit dans pred(Safe) ou bien on soit dans Safe
hence the resilience problem is decidable.
\end{proof}
